Taking Conditional Independence Seriously:\\A Latent Network Approach to Dyadic Analysis

### 2017-10-27 paper/4_app.tex
app sec changes

We choose five prominent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. These studies are recent and have been cited over 100 times. Each of these pieces was published in a prominent journal and is well-known in the literature. Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables. In addition, almost all of these studies adjusts the standard errors in an attempt to account for the clustering of observations. We follow the strategy of the original study in our re-estimation of the original models.

@@

We obtained the data for each of these studies from their replication archives and replicated the main results of each of the articles.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} We also examine each of the models using the AME framework described above.  Our goal is to ascertain whether the ignored interdependencies---the non-iid structure of the relational data---would result in different model estimates when they were addressed in an AME framework, and more importantly to see if there were substantive opportunities that were presented with the dynamic factor approach.  Finally, in each study we assess whether any new substantive findings emerge or if any previous findings disappear once the interdependencies in the data are modeled.

The broader goal, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that ignores dependencies.  Based in part on our results as well as the statistical characteristics of the AME framework, we believe that it does. The dynamic latent factor model provides a step forward in the modeling of international relations.

% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models. Generally, interpretations focus on a small set of independent variables.  For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework. Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.

@@

In addition, we also examine the predictions made with each approach using an out-of-sample cross validation strategy. By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures. To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models. To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
%\textcolor{red}{This is inaccurate for directed dyads. I think. MG} 
Then for each $s \in \{1,\ldots,k\}$, we:

@@

@@

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.  They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent. They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around.  They use the MID data, but note that:

\begin{quote}
	We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation. This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force. We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets. \ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
\end{quote}

Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero. In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one. Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies. Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework.\footnote{Further details on each of the regression parameter estimates from the replication can be found in the Appendix.}

One takeaway is that many studies make knowledge claims based on differences in statistical significance between covariates. These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.

@@

\citet{mcdonald:2004} studies whether trade promotes peace between nations. The link between conflict and trade is perplexing, with many persistent, yet competing explanations. \citet{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots (547)''. Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade. \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measures free trade in two ways. The first captures the idea that larger protected sectors generate greater societal pressures resulting in pockets of support for war. This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad. This measure captures the score of the state in the dyad that possesses higher barriers to trade (560). \citet{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP'' (560). The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad. \citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables. The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.

Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.  Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$. 
%Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict. 
This indicates that conflict is less likely between members of a trade community. Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.  The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$). At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).

@@

In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007,ward:etal:2013}.  Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. In the original, this negative linkages is not significant, but in the AME estimation this negative linkage is more precisely estimated. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.  One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. %Her time frame is from 1946-1999.  
The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.  These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340). She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). 
	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
	}
\end{figure}
\FloatBarrier

The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations. When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, and Israel who often send conflict to similar targets. There is a cluster of ``rogue states,'' who are receivers of conflict, in the top left (Iran, Iraq, Syria, Libya, North Korea). In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China. In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior. In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects. In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.  In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.

\begin{figure}
\centering   
	\subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
	\subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
\end{figure}
\FloatBarrier

>>>>

We choose five prominent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. These studies are recent and have been cited over 100 times. Each of these pieces was published in a prominent journal and is well-known in the literature. Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables. In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations. We follow the strategy of the original study in our re-estimation of the original models.

@@

We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data make much difference in what we have learned about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
Typically, scholars have a small set of independent variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models. For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework. Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}

@@

Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy. Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
%\textcolor{red}{This is inaccurate for directed dyads. I think. MG} 
Then for each $s \in \{1,\ldots,k\}$, we:

@@

Next we discuss, each of the replications in more detail and highlight the substantive insights that can be drawn from the AME framework.

@@

% They use the MID data, but note that:
% \begin{quote}
% 	We code a state as having initiated a dispute if it is on `side A' of a MID, the conventional approach to coding initiation. This means that the state was on the side that took the first action in the dispute, whether that action was the threat, display, or use of force. We code joiners as initiators or targets, though the results do not change if we do not code joiners as initiators or targets. \ldots Though coding initiation will always be difficult, the `side A' variable has been widely used in past conflict scholarship (page 334).
% \end{quote}

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes.  They use directed dyads and find that dyads involving a democratic leader on the one hand and a personalist dictator on the other tend to be violent. They also discover that dictators are likely to challenge democracies, but that this is not reciprocal.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, than the other way around. Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in these kinds of studies, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

In the original model, the variable ``Pers/Democ Directed Dyad" (which represents a Personalist $\rightarrow$ Democractic directed dyad) is clearly positive while the variable ``Democ/Personalist Directed Dyad'' is zero and the difference between the two coefficients is clearly distinct from zero. In our re-estimation using the AME framework, we also find that Pers-Democ directed dyad has a positive effect with zero excluded from the 95\% confidence interval while Democ-Pers directed dyad is indistinguishable from zero. Using this model, however, we can no longer conclusively say that the Pers/Democratic coefficient is larger than the Democ/Personalist one. Our re-estimation using the AME approach therefore cast some doubt on Reiter \& Stam's key claim that MIDs initiated by personalist dictatorships against democracies are more likely than MIDS initiated by democracies. Further, the effect of most of the covariates in the literature thought to predict interstate MIDs are much closer to zero when using the AME framework. An important takeaway here is that many studies make knowledge claims based on differences in statistical significance between covariates. These differences may change dramatically when interdependencies are taken into account directly, as they do in this study.

@@

\citet{mcdonald:2004} studies whether trade promotes peace between nations. The link between conflict and trade is perplexing, with many persistent, yet competing explanations. \citet[p. 547]{mcdonald:2004} includes the argument that interdependence between states ``makes conflict less likely because of its efficiency over conquest in acquiring resources\ldots''. Accordingly, his primary contribution is to provide evidence challenging the generalized linkage between peace and trade and to offer a new measurement of the key independent variable, trade. \citet{mcdonald:2004} refines the trade variable, arguing that \textit{free} trade, rather than trade alone, reduces the likelihood of conflict between states. His key hypothesis is that greater levels of protection increase the probability of interstate conflict, an argument that builds on the work of classic liberalism and connects free trade to the power of domestic audiences. \citet{mcdonald:2004} measures free trade in two ways. The first captures the idea that larger protected sectors generate greater societal pressures resulting in pockets of support for war. This protection variable measures the proportion of customs revenue divided by total imports in the state that possesses the greater such ratio in each dyad. This measure captures the score of the state in the dyad that possesses higher barriers to trade. \citet[p. 560]{mcdonald:2004} also includes a measure of economic integration  calculated as ``the lower proportion of total dyadic trade (imports plus exports) divided by state $i$'s GDP or total dyadic trade divided by state $j$'s GDP''. The binary, dependent variable is the onset of a new militarized interstate dispute within a given dyad. \citet{mcdonald:2004} employs logistic regression to examine the putative statistical significance of these variables. The models include splines to correct for temporal dependence, and robust standard errors clustered on each dyad.

Our re-estimation with AME reveals that trade relations are highly interdependent and exhibit important patterns of transitivity.  Or, in other words, if countries $i$ and $j$ are highly dependent and countries $j$ and $k$ are also highly dependent, then we are likely to observe high dependency between countries $i$ and $k$. This indicates that conflict is less likely between members of a trade community. Once we control for these dependencies, we can more clearly interpret the positive link between trade and conflict.  The most striking thing is that AME finds a positive conditional association between trade dependence and conflict ($\hat{\beta}= 18.4, \quad \sigma_{\hat{\beta}} = 28.6$), while the comparable numbers for the logistic regression found in the original articles are negative ($\hat{\beta}= -22.2, \quad \sigma_{\hat{\beta}} = 15.2$). At the same time, the AME has ROC and PR curves (shown in the Appendix) that dominate the results found in McDonald (2004).
% Our findings support those of Traag and Lupu (2013) which argue that indirect trade relations reduce the probability of conflict.

@@

\nocite{rose:2004}
In 2004, Andrew Rose published a study in the \textit{American Economic Review} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \citep{tomz:etal:2007,ward:etal:2013}.  Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion is that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO---as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar. The main result of the model---the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables---does not prevail in the AME model. In the original, this negative linkage is not significant, but in the AME estimation this negative linkage is more precisely estimated. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.  These insights are mainly emphasized in the paper by the parameter estimates depicted in Tables 1 and 2 (pages 339--340) from the paper. She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

\begin{figure}[!h]
	\includegraphics[width=\textwidth]{weeks_circPlot.pdf}
	\caption{\label{fig:weekscirc} Visualization of multiplicative effects for Weeks (2012). Each circle designates a country and the color corresponds to the legend at the center of the visualization. Countries that cluster together in the outer ring are those that were found by the model to have similar sending patterns, meaning that they tend to send conflict to similar sets of countries. The inner ring clusters countries by the similarity of who they receive conflict from. 
	%Blue represents groups with common sending patterns and red represents groups with common receiving patterns.
	}
\end{figure}
\FloatBarrier

The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the GLM results and those of the AME Model. In Weeks' initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone. When we examine the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence---in the top left corner we see states like the US, the UK, Germany, Canada and Israel. These states cluster together in the outer ring of this visualization because they tend to send conflicts to similar targets.\footnote{For details on how to interpret this type of visualization see \citet{minhas:etal:2016:arxiv}.} Conversely, in the bottom right of the outer ring, we see a cluster of authoritarian countries like Iraq, Russia, Syria, North Korea, and China. In the inner ring, the proximity of countries is determined by the degree to which they receive conflict from the same countries. In general, the clusters found on both the inner and outer rings have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category). In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type. Structural equivalence is present even when accounting for nodal characteristics like regime type.  The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects.

% In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out-of-sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.  In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.

### 2017-10-27 paper/6_appendix.tex
app sec changes

Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis. For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.

@@

% already show roc-pr in main text
% \begin{figure}
% 	\centering   
% 	\subfigure[AUC]{\label{fig:weeksroc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
% 	\subfigure[Precision and Recall]{\label{fig:weekspr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
% 	\caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves, PR curves, and separation plots.}
% \end{figure}
% \FloatBarrier
\clearpage

>>>>

Additionally, for each replication we provide a more detailed visualization illustrating the results of our out-of-sample performance analysis. %For Weeks (2012) we already did so in the main body of the text so we eschew from doing so again here.

@@

% already show roc-pr in main text
\begin{figure}
\centering   
  \subfigure[AUC]{\label{fig:weeksauc}\includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf}}
  \subfigure[Precision and Recall]{\label{fig:reitpr}\includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf}}
  \caption{Assessments of out-of-sample predictive performance for Weeks (2012) using ROC curves and PR curves}
\end{figure}
\FloatBarrier
\clearpage

### 2017-10-27 paper/netsMatter.tex
app sec changes

Much international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data. Yet, much contemporary research brushes aside the complications that come with analyzing relational data that are interdependent. Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches. As a result, many studies often produce results with biased coefficient estimates and poorly calibrated standard errors. These biases have profound consequences for evaluating parametric models. 
%parahere?
We present an alternative, regression-based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data. It uses a set of additive and multiplicative effects within a Bayesian regression framework. To highlight the utility of this framework for studying relational data, we present a simulation highlighting the ability of this approach to account for the dependencies that frequently characterize relational data such as often utilized in international relations. In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to the proposed latent factor approach. For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables and the hypotheses they derive from. The latent factor approach also dominates the current status quo approach in terms of out-of-sample cross-validations, rendering it more useful in hypothesis testing as well as forecasting. In short, it seems able to better capture the data generating process behind outcomes of interest.
%parahere?
We summarize the features of this latent factor model, and provide guideance on how to implement it in additional studies.
\end{abstract}
}}

@@

\newpage
% Bib stuff
\clearpage
%\bibliography{/Users/s7m/whistle/master}
\bibliography{/Users/mdw/Git/whistle/master}
%\bibliography{/Users/cassydorff/ProjectsGit/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}

>>>>

	Much of international relations scholarship concerns dyads: dyadic hypotheses and propositions, and especially dyadic data. Yet, contemporary research brushes aside the complications that come with analyzing relational data that are interdependent. Dyadic observations do not typically satisfy the conditional independence criterion required for many statistical approaches. As a result, many studies often produce results with biased coefficient estimates and poorly calibrated standard errors. These biases have profound consequences for evaluating parametric models. We present an alternative, regression-based approach that accounts for the dependencies that complicate inferential analysis in the context of relational data. To highlight the utility of this framework for studying relational data, we first present a simulation exercise highlighting the model's ability to account for the dependencies that emerge in relational data. In addition, we replicate five prominent studies in recent international relations scholarship, comparing the standard approach to our alternative. For each study, we find that conventional methods overstate the effect of key variables in the study of interstate conflict (and trade), underestimate the uncertainty in these effects, and in some cases lead researchers to faulty conclusions about the statistical significance and substantive importance of their variables. Further, we show that our approach dominates in terms of out-of-sample cross-validations, rendering it more useful in forecasting applications and in modeling the data generating process behind outcomes of interest.
\end{abstract}
}}

@@

\newpage
% Bib stuff
\clearpage
\bibliography{/Users/s7m/whistle/master}
% \bibliography{/Users/mdw/Git/whistle/master}
%\bibliography{/Users/cassydorff/ProjectsGit/master}
% \bibliographystyle{authordate1}
\bibliographystyle{sm}
% \clearpage
% \section*{All About Us}
